A Plasma Torus Around a Young Low-Mass Star

### 2025-04-04 paper/ms.tex
update

\begin{figure}[!tp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/sf6.pdf}
  \caption{Time-variable fit to H$\alpha$ line profiles.
  {\it Left column:} Raw spectrum at each epoch $f_{\lambda}$ minus
  the time-averaged spectrum $f_{\langle t \rangle}$ (as in
  Figure~\ref{fig:spec}e).
  Underplotted sinusoids are not fits; they are meant to guide the eye.
  {\it Middle columns:} Model of emission from the inner clump (sum of
  two gaussians) and the outer clump (single gaussian), plotted over
  the data.
  {\it Right column:} Residual of the left column after subtracting
  the sum of the two middle columns, leaving variability in the line
  core.  
  Section~\ref{subsec:halpha} discusses the use of this model.
  }
  \label{fig:halphamodel}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/sf7.pdf}
  \caption{Orbits fit to mean radial velocities (RVs) extracted from 
  H$\alpha$ profile fits in Figure~\ref{fig:halphamodel}.
  Each marker represents the measured best-fit mean at a given epoch.
  Solid circles were used in the fits; transparent X markers
  were excluded due to concerns about their epoch-by-epoch
  interpretation (see text).  The purpose of this exercise was to
  quantify the orbital periods and semi-amplitudes of the
  circumstellar clumps.  The inner double-peaked clump is shown in
  orange and blue; the outer clump is shown in green.  Five draws from
  each posterior probability distribution are plotted.}
  \label{fig:orbits}
\end{figure}

Taking these caveats into account, we used the results from Extended
Data Table~\ref{tab:halphamodelparams} to evaluate various models for
the orbits.  Our main goal in this exercise was to quantify the
orbital periods and velocity semi-amplitudes of the clumps.
Figure~\ref{fig:orbits} shows representative results for the
circular orbit case; we found no reason to extend the exercise to
eccentric orbits.
Based on visual inspection of the line profile fits, we manually excluded 
epochs where we do not believe our gaussian profile fitting provided a
reliable representation of the circumstellar emission.  We then performed
the orbit fits using NumPyro to sample over the Gaussian likelihood with the
NUTS algorithm \cite{Phan2019}.  When performing the fits, we adopted
the uncertainties on each estimated mean radial velocity value, and
added a jitter term in quadrature to inflate these nominal measurement
uncertainties.  This procedure yielded period and semi-amplitude
values for circular orbits of
$P_0$=$3.92\pm0.03$\,hr, $K_0/v_{\rm eq}$=2.07$\pm$0.04;
$P_1$=$3.92\pm0.06$\,hr, $K_1/v_{\rm eq}$=2.88$\pm$0.10;
and $P_2$=$3.35\pm0.35$\,hr, $K_1/v_{\rm eq}$=4.16$\pm$0.56.
The periods for the inner double-peaked clump are consistent
with the photometric 3.930$\pm$0.001\,hr period to a precision of
about two minutes.
The ``period'' for the outer clump is more ambiguous because the
H$\alpha$ data only support the idea of a periodic orbit of material
well-fit by gaussian emission from $\phi$$\approx$0.5-1.0.
From $\phi$=0-0.2, there is no detectable emission,
and from $\phi$=0.2-0.4, the suddenly-detectable emission spans
1-4\,$v_{\rm eq}$ without taking a clear gaussian shape.
While the outer-most edge of the $\phi$=0.2-0.4 emission provides a
closer match to the expectation of a circular orbit, 
the idea of invoking a particular functional form for this component 
seems fine-tuned.
We instead emphasize that although this emission is present, 
its variability in time is poorly explained by the idea of a
stable, long-lived clump of material.
Although additional observations would be needed to conclusively determine whether or not 
this component of the system
is long-lived, the lack of any photometric variability in the TESS
data during its expected transits might suggest that it is not.

{\it Constraints on physical dimensions of emitting region}---
%TODO FIXME TODO FIXME

IMPLICATIONS:
The secondary eclipse of the `inner clump' at $\phi$$\approx$0.5 is not
associated with a broadband flux dip.   The clump does not therefore
contribute significant broadband emissivity across the entire TESS
band; only in H$\alpha$.

>>>>

\begin{figure}[!tp]
  \centering
  \includegraphics[width=0.55\textwidth]{figures/sf6.pdf}
  \caption{Time-variable fit to H$\alpha$ line profiles.
  {\it Left column:} Raw spectrum at each epoch $f_{\lambda}$ minus
  the time-averaged spectrum $f_{\langle t \rangle}$ (as in
  Figure~\ref{fig:spec}e).
  Underplotted sinusoids are not fits; they are meant to guide the eye.
  {\it Middle columns:} Model of emission from the inner clump (sum of
  two gaussians) and the outer clump (single gaussian), plotted over
  the data.
  {\it Right column:} Residual of the left column after subtracting
  the sum of the two middle columns, leaving variability in the line
  core.  
  Section~\ref{subsec:halpha} discusses the use of this model.
  }
  \label{fig:halphamodel}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/sf7.pdf}
  \caption{Orbits fit to mean radial velocities (RVs) extracted from 
  H$\alpha$ profile fits in Figure~\ref{fig:halphamodel}.
  Radial velocities on the vertical axis are in units of the
  equatorial velocity, $v_{\rm eq}$=130\,\kms.
  Each marker denotes the best-fit mean at a given epoch.
  Solid circles were adopted in the fits; transparent X markers
  were excluded due to concerns about their reliability
  (see text).  The inner double-peaked clump is shown in
  orange and blue; the outer clump is shown in green.  Five draws from
  each posterior probability distribution are shown.}
  \label{fig:orbits}
\end{figure}

Accounting for these caveats, we used the results from Extended
Data Table~\ref{tab:halphamodelparams} to evaluate various models for
the orbits.  Our main goal was to quantify the orbital periods and
velocity semi-amplitudes of the clumps.
Figure~\ref{fig:orbits} shows representative results for the case of a circular
orbit; eccentric orbits yielded similar results.  Based on visual
inspection of Extended Data Figure~\ref{fig:halphamodel}, we manually excluded
epochs when our gaussian profile fitting failed to either detect or
else adequately represent the circumstellar emission.  We then used
NumPyro to sample the Gaussian likelihood for a Keplerian orbit
with the NUTS algorithm \cite{Phan2019}.  We used the measurement
uncertainties from each estimated mean radial velocity value and
included an additional jitter term in quadrature.  This procedure
yielded orbital periods and semi-amplitudes of
$P_0$=$3.92\pm0.03$\,hr, $K_0/v_{\rm eq}$=2.07$\pm$0.04;
$P_1$=$3.92\pm0.06$\,hr, $K_1/v_{\rm eq}$=2.88$\pm$0.10;
and $P_2$=$3.35\pm0.35$\,hr, $K_1/v_{\rm eq}$=4.16$\pm$0.56.
The periods for the inner double-peaked clump are consistent with the
photometric $3.930\pm0.001$\,hr period within a precision of two
minutes.  The ``period'' for the outer clump is ambiguous because the
H$\alpha$ data only support the idea of a periodic orbit of material
well-fit by gaussian emission from $\phi$$\approx$0.5-1.0.  From
$\phi$=0-0.2, there is no detectable emission, and from
$\phi$=0.2-0.4, the emission spans 1-4\,$v_{\rm eq}$ without taking a
clear gaussian shape.  While the outer-most edge of this
$\phi$=0.2-0.4 emission provides a plausible match to the expectation
of a circular orbit, the idea of invoking a particular functional form
for this component seems fine-tuned.  We instead emphasize that
although this emission is present, its variability in time is poorly
explained by the idea of a stable, long-lived clump of material.
Although additional observations would be needed to conclusively
determine whether or not this component of the system is long-lived,
the lack of any photometric variability in the TESS data during its
expected transits suggests that it is not.

{\it The physical dimensions of emitting region}---The measured
velocity widths of the emitting circumstellar material contain
information about the size of the emitting region via the condition
for rigid corotation.
Consider a clump in cylindrical coordinates with arbitrary radial
extent $r$, azimuthal extent $\ell$, and height $z$.
At quadrature, the observed velocity width is sensitive to the radial
extent of this material because it is being viewed face-on.
During transit, the observed velocity width is sensitive to the
azimuthal extent.
Although the vertical height is less constrained for an edge-on system, 
a rough estimate can follow from assuming that the depth of
the broadband TESS absorption $\delta$ should scale with the product
$\ell z$.

At quadrature, the inner clumps show $\sigma_i$$\approx$0.24\,$v_{\rm
eq}$, implying that 68\% of the emission comes from a volume with
radial extent $r_i$=$2\sigma_i / \Omega$$=$0.48\,$R_\star$, and
that 95\% of the emission comes from within 0.96\,$R_\star$.
These values have relative uncertainties of $\approx$5\%, based on the
uncertainties in the measured velocity
dispersions.  The two inner clumps are centered at distances of
2.07\,$R_\star$ and 2.88\,$R_\star$.  There is therefore 
physical overlap in their spatial distributions.  These
two clumps could in fact be a single clump with an optically thick H$\alpha$
line core.
Regardless of this nuance, the implication is that the combined 
extent in the radial dimension of these two inner emitting clumps is
roughly equal to the star's diameter.
At mid-transit, these inner clumps seem to have a similar velocity
dispersion, although with greater uncertainty due to the differences
between the $\phi$=0 and $\phi$=1 transits.  This nonetheless suggests
a 1$\sigma$ emission contour in the azimuthal dimension of
$\ell$$\approx$0.5\,$R_\star$.

%TODO FIXME ADD DISCUSSION OF VERTICAL DIMENSION...
%TODO FIXME TODO FIXME
